---
title: "ST512-SP26-Homework-5"
subtitle: Jason Menjiar-Hernandez
format: pdf
---

## Problem 1

**a**

| Source              | df  | SS   | MS    | F-value | p-value  |
|---------------------|-----|------|-------|---------|----------|
| **Factor A**        | 4   | 375  | 93.75 | 27.79   | 3.66e-14 |
| **Factor B**        | 2   | 462  | 231.0 | 68.49   | 0        |
| **AxB Interaction** | 8   | 38   | 4.75  | 1.40    | 0.210    |
| **Error**           | 75  | 253  | 3.373 | --      | --       |
| **Total**           | 89  | 1128 | --    | --      | --       |

**b**

Based on the ANOVA table, the interaction factor has a p-value of 0.210. Using a 0.05 significance level, we fail to reject the null hypothesis. Therefore, there is NOT a significant interaction between factor A and factor B.

**c**

Based on the ANOVA table, for Factor A the p-value was calculated to be 3.66e-14. Using a significance value of 0.05, we reject the null hypothesis. Therefore, there is a significant difference in mean response across the levels of factor A.

**d**

Based on the ANOVA table, for Factor B the p-value was calculated to be 0. Using a significance value of 0.05, we reject the null hypothesis. Therefore, there is a significant difference in mean response across the levels of factor B.

## Problem 2

**a**

a=4, b=3, n=3, N=36, SSE=720

df (error)= 24, MSE=720/24=30.0

No interaction --\> each factor A mean averages over b\*n=9

Tukey MSD:

$q_{\alpha,4,24} * \sqrt{MSE/n_A)}$ = 3.90 \* $\sqrt{30/9}$ = 7.12

Bonferroni MSD:

$t_{\frac{\alpha}{2C},24}$ \* $\sqrt{2MSE/n_A}$ = 2.875 \* $\sqrt{60/9}$ = 7.42

**b**

Interaction --\> compare Factor A within one level of Factor B, n=3

Tukey MSD:

$q_{\alpha,4,24} * \sqrt{MSE/n_A)}$ = 3.90 \* $\sqrt{30/3}$ = 12.33

Bonferroni MSD:

$t_{\frac{\alpha}{2C},24}$ \* $\sqrt{2MSE/n_A}$ = 2.875 \* $\sqrt{60/3}$ = 12.86

## Problem 3 



## R code

**problem 1**

```{r}
1-pf(27.79,4,75)
1-pf(68.49, 2,75)
1-pf(1.40,8,75)
```

**problem 2**

```{r}
qtukey(.95,4,24)
qt(0.0041667,24)
```
